Asymmetric relationships and their impact on coexistence

### 2023-01-26 paper/Paper.Rnw
small change that should have been included in the previous commit
edited Theory section and caption of fig 2 and corresponding sections of sup mat

We here pursue goal G1 from the Introduction: to augment MCT to illuminate how ATA
plays a role in coexistence. We develop each step of our new theory for a general 
two-species model (below), and also specifically for the lottery model of the Introduction 
(SI section \ref{SI-SIsect:TheoryForLottery}).
MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
quantify the contributions of 
multiple mechanisms to coexistence. But we show below
that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
So we begin by defining storage effects. Our mathematical notation parallels
that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$. We also assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
presentation of new ideas.

The storage effect for species $i$ is the contribution of covariation between $E$ and 
$C$ to a difference between rare and common species mean growth rates, when $i$ is rare.
Specifically, storage effects for species $i$ are 
\begin{equation}
\Delta_i^{(EC)} = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
\end{equation}
where: $\olr_{i \bs i}$ refers to the mean ``growth when rare'' [GWR] of species $i$, that is, the mean growth rate of 
species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
$\olr_{j \bs i}$ refers to the mean steady-state 
growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
influences of environment and competition on either species have been rendered independent.  
A subscript 
$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
The \emph{scaling factor}
$q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two
alternative definitions of $q_{ij}$ that we use. 
Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions.
Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
An expression for $\olrs_{i \bs i}$ is obtained from $\olr_{i \bs i}$ by replacing the bivariate random 
variable $(E_i,C_{i \bs i})$ by a random variable with the same marginal distributions, but independent components, 
$(E_i^\shp,C_{i \bs i}^\shp)$, so $\olrs_{i \bs i}=\E[r_i(E_i^\shp,C_{i \bs i}^\shp)]$.
Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.
Though some readers may find it slightly confusing to simultaneously use the variable 
$E_i$ and the operator $\E(\cdot)$, both of these notational choices are standard, so it would 
be more confusing to use alternative notation. Like all variables, $E_i$ is rendered in italic; like
all operators, $\E(\cdot)$ is rendered upright.

We now specify how to decompose storage effects into
contributions due to ATAs and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between rare and common species mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$.
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
correlation \emph{per se}, remaining after having removed the effects of ATAs.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
The new symbol should correspondingly be pronounced ``partial sharp''.
The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
below for additional details.

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal.  Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$. We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$. The value of $\ddot{\rho}_i$ is then determined 
in one of 
four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.

The other mechanisms in MCT that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas ATAs are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
coexistence operating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of MCT which was goal G1 of the Introduction. One typically (but see below) must examine this quantity for each species, $i$,
to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms.

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, and Robin Snyder for helpful discussions. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, and the California Department of Fish and
Wildlife Delta Science Program. 
%***DAN: DAN: if we take Ellner off the author list add him back here, and also add him back to the acks in the README. Also take him off the author list in sup mat.

%you store your bibliographic info in refs.bib, see that file
\bibliographystyle{ecology_letters2}
\bibliography{refs}

@@

\begin{figure}
\includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
%***DAN: I edited the caption (2023 01 25) but need to replace the figure once I know where in the repo 
%structure it is generated and stored, and after I have re-run Jasmin's code
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (g,h)]. 
The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
pictured and then transforming the marginals as specified in Methods.
(d-f) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
respectively corresponding variables (a-c; see Introduction).
(g,h) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (g) showing 
left-tail association and (h) showing right-tail association. \label{fig:pedagog1}}
\end{figure}

\begin{figure}
\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
\caption{Our simulated lottery-model populations influenced by asymmetric tail associations (ATAs) recovered 
from low densities less often and 
spent more time in a state of one-species dominance, demonstrating that ATAs can influence aspects of 
species coexistence. Panel (a) shows species 1 from the lottery 
model subject to environmental noise with symmetric tail associations (Fig. \ref{fig:noise}b). Panel (b) shows species 1 from the 
same model with the same parameters, but subject to noise with left-tail association (Fig. \ref{fig:noise}a).
Both simulations were run for \Sexpr{M} time steps, though only the first $1000$ are shown. Periods of 
``noticeable coexistence'' of the species, defined to be periods when each species comprises at least $5\%$ of 
total abundance, are marked in grey. The average duration of periods of noticeable coexistence 
was \Sexpr{fig2$coMets$Sym$co} time steps for panel (a) and \Sexpr{fig2$coMets$Asy$co} for panel (b), over the full length of the simulations; 
whereas the average duration of periods of
one-species dominance was \Sexpr{fig2$coMets$Sym$dom} time steps for panel (a) and \Sexpr{fig2$coMets$Asy$dom} for (b). Thus, while periods of 
noticeable coexistence were of similar duration for the two scenarios, periods of one-species dominance were much longer,
on average, for the ATA case. Parameters were $\mu_1 = \Sexpr{fig2$params[1]}$,
$\mu_2 = \Sexpr{fig2$params[2]}$, $\sigma = \Sexpr{fig2$params[3]}$, and $\delta = \Sexpr{fig2$params[4]}$. 
See Introduction and Fig. \ref{fig:noise} caption for model details. \label{fig:pedagog}}
\end{figure}

>>>>

We here pursue goal G1 from the Introduction: to augment MCT to show formally how ATAs play a role in coexistence. 
We develop our new theory for a general two-species model, later applying it to specific scenarios.
MCT and its recent computational extensions \citep{Ellner_2016,Ellner_2019} 
quantify the contributions of 
multiple mechanisms to coexistence. But we show below
that, for purely temporal variation, only one mechanism relates to tail associations: storage effects. 
So we begin by defining storage effects. Our mathematical notation parallels
that of \cite{Ellner_2016,Ellner_2019}. It is assumed that the growth rate $r_i(t)=\ln(N_i(t+1)/N_i(t))$ 
of species $i$ can be written as an increasing function of
an environment-dependent factor $E_i(t)$ and as a decreasing function of a quantity representing
competitive pressure on species $i$, $C_i(t)$. We also assume, for simplicity, 
that the $E_i(t)$ are iid through time, and likewise for the
$C_i(t)$. Though the iid assumption can be circumvented, we make the assumption because it simplifies the
presentation of new ideas.

\subsection{The storage effect}

The storage effect for species $i$ is the contribution of covariation between $E$ and 
$C$ to a difference between rare and common species mean growth rates, when $i$ is rare.
Specifically, storage effects for species $i$ are 
\begin{equation}
\Delta_i^{(EC)} = (\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i}),\label{eq:storage_effects_gen}
\end{equation}
where: $\olr_{i \bs i}$ refers to the mean ``growth when rare'' [GWR] of species $i$, that is, the mean growth rate of 
species $i$ when it is at negligible abundance and species $j$ (the other species) is at stochastic steady state;
$\olr_{j \bs i}$ refers to the mean steady-state 
growth rate of species $j$ when species $i$ is absent or at negligible abundance (this is $0$, \emph{a priori});
and $\olrs_{i \bs i}$ and $\olrs_{j \bs i}$ are analogous quantities, but for an alternative modelling scenario, identical except that
influences of environment and competition on either species have been rendered independent.  
A subscript 
$i \bs i$ always indicates a quantity computed for species $i$ when that species is at negligible abundance, and 
a subscript $j \bs i$ indicates a quantity computed for $j$ when $i$ is absent.
The \emph{scaling factor}
$q_{ij}$ measures the relative sensitivities to competition of the two species, see \cite{Chesson_1994},
\cite{Ellner_2016} and SI section \ref{SI-SIsect:scaling_factors} for the two
alternative definitions of $q_{ij}$ that we use. 
Since it was assumed that $r_i=r_i(E_i,C_i)$, we can write $\olr_{i \bs i}=\E[r_i(E_i,C_{i \bs i})]$, where
$\E(\cdot)$ denotes expected value just as does an overbar, but we use $\E$ for longer expressions.
Also, $\olr_{j \bs i}=\E[r_j(E_j,C_{j \bs i})]$.
An expression for $\olrs_{i \bs i}$ is obtained from $\olr_{i \bs i}$ by replacing the bivariate random 
variable $(E_i,C_{i \bs i})$ by a random variable with the same marginal distributions, but independent components, 
$(E_i^\shp,C_{i \bs i}^\shp)$, so $\olrs_{i \bs i}=\E[r_i(E_i^\shp,C_{i \bs i}^\shp)]$.
Likewise, $\olrs_{j \bs i}=\E[r_j(E_j^\shp,C_{j \bs i}^\shp)]$.

\subsection{Decomposing the storage effect}

We now specify how to decompose storage effects into
contributions due to ATAs and due
to $EC$ correlation \emph{per se}. 
Storage effects for species $i$ were computed by comparing
differences between rare and common species mean growth rates for the original model and for a 
modified model for which the components of $(E_i,C_{i \bs i})$ and those of $(E_j,C_{j \bs i})$
were rendered independent of each other [see (\ref{eq:storage_effects_gen})]. 
Because these components were rendered 
completely independent in the modified model, storage effects quantify the contribution of the 
whole $EC$ relationship, in all its aspects, to the rare/common species growth rate
difference. But we can instead consider distributions $(E_i^\nat,C_{i \bs i}^\nat)$ and 
$(E_j^\nat,C_{j \bs i}^\nat)$ with the same
marginal distributions, respectively, as $(E_i,C_{i \bs i})$ and $(E_j,C_{j \bs i})$,
and with some but not all aspects of the relationship 
between the components altered relative to the original distributions. Specifically, asymmetries of tail association are removed -- see below for how this is done. We then define 
$\olrn_{i \bs i}$ and $\olrn_{j \bs i}$ based on a modified model using  $(E_i^\nat,C_{i \bs i}^\nat)$
and $(E_j^\nat,C_{j \bs i}^\nat)$, and we consider the decomposition 
\begin{align}
\Delta_i^{(EC)} &= [(\olr_{i\bs i} - q_{ij} \olr_{j \bs i})-(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})]+
[(\olrn_{i \bs i}-q_{ij} \olrn_{j \bs i})-(\olrs_{i \bs i}-q_{ij} \olrs_{j \bs i})] \label{eq:storage_effects_further1}\\
&= \Delta_i^{[EC]} + \Delta_i^{[E \nat C]}.\label{eq:storage_effects_further}
\end{align}
The term $\Delta_i^{[EC]}$ is the component of storage effects due to 
ATAs between $E_i$ and $C_{i \bs i}$, and between $E_j$ and $C_{j \bs i}$. We will refer to this term as 
the \emph{ATA effect on coexistence}. 
The term $\Delta_i^{[E \nat C]}$ is the component of storage effects due to $EC$ 
correlation \emph{per se}, remaining after having removed the effects of ATAs.
The notation $\nat$ was chosen because this new symbol is part of the symbol $\shp$ (i.e., two of its four lines), 
corresponding to the removal of part of the relationship between $E$ and $C$ components (namely, ATAs).
The new symbol should correspondingly be pronounced ``partial sharp''.
The brackets in $\Delta_i^{[E \nat C]}$ and $\Delta_i^{[EC]}$ are intended to indicate that 
(\ref{eq:storage_effects_further}) is a decomposition within a decomposition, i.e., (\ref{eq:storage_effects_further}) is a decomposition
of storage effects, which are themselves a term in the MCT decomposition \citep{Ellner_2019} -- see 
below for additional details.

\subsection{Removing asymmetries of association}

How are asymmetries of tail association removed while leaving $E$ and $C$ marginal distributions and overall 
correlation unchanged? Fig. \ref{fig:pedagog2} provides an illustration of the steps we describe here. 
We assume, for simplicity, that the cumulative distribution functions (CDFs) of $E_i$ and 
$C_{i \bs i}$, which we denote $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively, are strictly monotonic and continuous. 
Letting $\varphi$ be the CDF of a standard normal distribution, we consider the random variables
$\varphi^{-1} \circ F_{E_i} (E_i)$ and $\varphi^{-1} \circ F_{C_{i \bs i}}(C_{i \bs i})$, which are standard normally
distributed, although they are not necessarily jointly bivariate normal (Fig. \ref{fig:pedagog2}a,b).  
Let $\rho_i$ be the covariance of
these random variables. We then define a bivariate normal random variable $(e_i,c_{i \bs i})$ such that
$\E(e_i)=\E(c_{i \bs i})=0$ and $\var(e_i)=\var(c_{i \bs i})=1$, and we denote 
$\cov(e_i,c_{i \bs i})$ by $\ddot{\rho}_i$ (Fig. \ref{fig:pedagog2}c). We define 
$E_i^\nat = F_{E_i}^{-1} \circ \varphi(e_i)$
and $C_{i \bs i}^\nat=F_{C_{i \bs i}}^{-1} \circ \varphi(c_{i \bs i})$ (Fig. \ref{fig:pedagog2}d). 
The value of $\ddot{\rho}_i$ is then determined in one of 
four alternative ways: $\ddot{\rho}_i = \rho_i$; or so that the Pearson, Spearman or Kendall correlation of 
$E_i^\nat$ and $C_{i \bs i}^\nat$ matches that of $E_i$ and $C_{i \bs i}$. The choice of bivariate normal 
$(e_i,c_{i \bs i})$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are symmetrically tail associated.
For those familiar with copula statistics (see, e.g., \citealp{Genest_2007} and \citealp{Ghosh_2020_AER}), 
we replaced the copula of $(E_i,C_{i \bs i})$ with a normal copula to form $(E_i^\nat,C_{i \bs i}^\nat)$.
The choice of $\ddot{\rho}_i$ ensures that $E_i^\nat$ and $C_{i \bs i}^\nat$ are correlated to the same extent as 
$E_i$ and $C_{i \bs i}$, according to whichever definition of correlation is considered preferable. Thus $E_i^\nat$ and 
$C_{i \bs i}^\nat$ have had their tail association rendered symmetric while retaining their correlation \emph{per se},
as desired. The random variable $(E_j^\nat, C_{j \bs i}^\nat)$ is created similarly.

\subsection{Our decomposition as an extention of modern coexistence theory}

The other mechanisms in MCT that apply in our modelling context of purely temporal variation
make no reference to covariation between quantities, whereas ATAs are 
defined in a covariation context. Therefore, storage effects are the only mechanism of coexistence
for which ATAs are relevant. Mathematical details supporting this claim are in SI section 
\ref{SI-SIsect:other_mechanisms}, but we here provide a conceptual understanding. \cite{Ellner_2016} and
\cite{Ellner_2019} provide a decomposition of the coexistence metric $\olr_{i \bs i} -q_{ij}\olr_{j \bs i}$,
which equals GWR of species $i$. The decomposition, which we use below, is similar to and motivated by the 
original decomposition of \cite{Chesson_1994}; the decomposition is how the mechanisms of 
coexistence operating in a given case are illuminated. The decomposition is 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{(EC)},\label{eq:previous_decomp}
\end{equation}
where: $\Delta_i^0$ is a baseline;
$\Delta_i^E$ is a contribution to coexistence of environmental variation;
$\Delta_i^C$ is a contribution to coexistence of variation in competition;
$\Delta_i^{(E\shp C)}$ is a contribution to coexistence of having simultaneous variation in both environment and competition, but not including the effects 
of covariation in these quantities; and
$\Delta_i^{(EC)}$ is a contribution to coexistence of $EC$ covariation itself -- storage effects. 
Only storage effects involve $EC$ covariation,
so the only mechanism of coexistence influenced by ATA between $E$ and $C$ is storage effects. Combining our decomposition 
(\ref{eq:storage_effects_further}) with the decomposition (\ref{eq:previous_decomp}) of \cite{Ellner_2019} gives 
\begin{equation}
\olr_{i \bs i} -q_{ij}\olr_{j \bs i}=\Delta_i^0+\Delta_i^E+\Delta_i^C+\Delta_i^{(E\shp C)}+\Delta_i^{[E \nat C]} + \Delta_i^{[EC]},\label{eq:full_decomp}
\end{equation}
which is the extension of MCT which was goal G1 of the Introduction. One typically (but see below) must examine this quantity for each species, $i$,
to obtain complete information about coexistence. To pursue goal G2, below, we will quantify
the magnitude of the new ATA term, $\Delta_i^{[EC]}$, relative to the magnitudes of the other terms for the 
lottery model described in the Introduction.

@@

The authors thank Shyamolina Ghosh for suggesting the research problem, and Lawrence Sheppard, Max Castorani,
Jon Walter, Jude Kastens, Vadim Karatayev, Emily Arsenault, Neil Thimmesch, and Robin Snyder for helpful discussions. 
The authors especially thank Steve Ellner for key advice at several watershed moments in the devlopment of the project. 
The authors were partly supported by US National
Science Foundation grants 1714195 and 2023474, the James S McDonnell Foundation, the California Department of Fish and
Wildlife Delta Science Program, and the Alexander von Humboldt Foundation.

\bibliographystyle{ecology_letters2}
\bibliography{refs}

@@

%***DAN: I edited the caption (2023 01 25) but need to replace the figure once I know where in the repo 
%structure it is generated and stored, and after I have re-run Jasmin's code
\begin{figure}
\includegraphics[width=.25\textwidth]{../results_figs/fig1_vertical.pdf}
\caption{Examples motivating our study of the effects of asymmetric tail associations (ATAs) on competition
and coexistence. (a-c) Bivariate random variables with standard normal marginals showing
alternative patterns of association in the left and right parts of the distributions. 
The two variables in (a) are more strongly related to each
other in the left parts of the distributions, termed ``left-tail association;'' those in (b) 
are symmetrically associated; and those in (c) 
are more strongly related to each other in the right parts, termed ``right-tail association''. All three cases have 
the same Pearson correlation, $\rho$, up to sampling variation, so Pearson correlation does not identify these 
very different patterns of association. For (a, c), we used asymmetrically tail associated cases which are perfectly 
correlated in the lower or upper halves of the distributions, though it is also possible to generate asymmetrically 
tail associated noise with imperfect correlations in both halves of the distributions [\citealp{Nelsen2006,Ghosh_2020_AER};
see also (g,h)]. 
The variables $(B_1,B_2)$ described in the Introduction and Methods were obtained by generating noise as 
pictured and then transforming the marginals as specified in Methods.
(d-f) Lottery model population simulations using $(B_1,B_2)$ obtained by transforming the 
respectively corresponding variables (a-c; see Introduction).
(g,h) Two examples of ATAs found in nature. Axes are population densities of two plankton species, 
\emph{Ceratium fusus} and \emph{Ceratium furca}, sampled in successive years in two of the 15 locations 
in the North Sea considered by \citep{Ghosh_2020_ecolevol}. ATAs differed by location, with (g) showing 
left-tail association and (h) showing right-tail association. \label{fig:pedagog1}}
\end{figure}

%***DAN: I edited the caption (2023 01 26) but this is the completely the wrong figure. I need to replace it 
%with Jasmin's new pedagogical figure once I know where in the repo 
%structure that one is generated and stored, and after I have re-run Jasmin's code, and after she has made the
%changes I recently asked for (see next comment).
%***DAN: I asked Jasmin to replace the \phi with \varphi in the axis labels on b and d, and to make the marginals
%on b and c be *standard* normal instead of just normal.
\begin{figure}
\includegraphics[width=.75\textwidth]{../results_figs/fig2.pdf}
\caption{Illustration of how asymmetric tail associations are removed while retaining marginal distributions and correlation 
between variables (see Theory). (a) An example original bivariate distribution $(E_i,C_{i \bs i})$, with ATA present. 
(b) Marginals are now normalized. The variables $E_i$ and $C_{i \bs i}$ are individually transformed to have standard normal 
marginals through a composition of the inverse of the standard normal cumulative distribution function (CDF), 
$\varphi^{-1}$, and 
each variable's CDF, $F_{E_i}$ and $F_{C_{i \bs i}}$, respectively.  
(c) A newly generated bivariate normal distribution, $(e_i,c_{i \bs i})$, with correlation $\ddot{\rho}=\rho$
selected to match that of the original distribution (see Theory). (d) Two new variables, 
$E_i^{||}$ and  $C_{i \bs i}^{||}$,
%***DAN: I replaced \nat in the above two expressions by || because for some reason the \nat does not
%compile when it occurs within a caption. Not sure how to deal with that, but leaving it alone for now.
which have the same marginals and correlation as the original $E_i$ and $C_{i \bs i}$ variables of (a) but 
which have symmetric association structure. These variables are obtained by transforming the marginals from (c), 
$e_i$ and $c_{i \bs i}$, using the inverses of the functions used in (b).
For illustration purposes we started with beta distributed $E_i$ and $C_{i \bs i}$. Contour lines are 
$\log_{10}$ of probability density on all panels. 
\label{fig:pedagog2}}
\end{figure}
